---
title: "A guide to importing Stata .dta files into R"
author: "Chris Penfold"
date: "31 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is intended as a step-by-step guide for importing large Stata `.dta` files, converting into R friendly formats, and then saving out as `.fst` files to allow quick random access to the file (without having to load the whole file). It also covers importing these files and retrieving the lost variable labels.

```{r cars}
library(haven)
library(webuse)
library(dplyr)
library(tidyverse)
library(fst)

census <- webuse("census") # use 'census' Stata dataset
census2 <- as_factor(census) # use haven to convert labelled variables into regular R factor variables

# save out an fst version for demo purposes
write.fst(census2, "I:/census2.fst")

census %>% select(state:pop65p) %>%  summary()
```

## Including Plots

These filter options are equivalent (truncated range of variables shown)

```{r }
census_fst <- read.fst("I:/census2.fst", as.data.table = FALSE)
census_fst %>% filter(region==levels(region)[1]) %>% select(state:pop65p)

census_fst %>% filter(region  %in% "NE") %>% select(state:pop65p)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
