---
title: "A guide to importing Stata .dta files into R"
author: "Chris Penfold"
date: "31 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is intended as a step-by-step guide for importing large Stata `.dta` files, converting into R friendly formats, and then saving out as either:

1.  `.fst` files

    1.  To allow quick random access to the file (without having to load the whole file).

    2.  It also covers importing these files and retrieving the lost variable labels.

2.  `.qs` files

    1.  A much quicker version of `saveRDS`

    2.  All attributes are preserved, but random access is not possible

```{r cars}
library(haven)
library(webuse)
library(dplyr)
library(tidyverse)
library(fst)

census <- webuse("census") # use 'census' Stata dataset
census2 <- as_factor(census) # use haven to convert labelled variables into regular R factor variables

# save out an fst version for demo purposes
write.fst(census2, "I:/census2.fst")

census %>% select(state:pop65p) %>%  summary()
```

## Including Plots

These filter options are equivalent (truncated range of variables shown)

```{r }
census_fst <- read.fst("I:/census2.fst", as.data.table = FALSE)
census_fst %>% filter(region==levels(region)[1]) %>% select(state:pop65p)

census_fst %>% filter(region  %in% "NE") %>% select(state:pop65p)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
